Was Devon Allen Unjustly Disqualified at the 2022 World Track and Field Championships?

### 2023-03-28 Manuscript/manuscript.tex
Meeting notes from 3/28 are in-line in the manuscript

\begin{abstract}
  At the 2022 World Track Championships in Eugene Oregon, Devon Allen was disqualified in the
  Men's 110 meter hurdle final after registering a reaction time of 0.099 seconds, 0.001 
  seconds faster than what is allowed.  Following the games, bloggers on the running website 
  LetsRun scrutinized the results and performed non-statistical analysis of the data.
  They found that the reaction time data from the 2022 World Track Championships seemed to be faster 
  compared to the other datasets they looked at, but did not perform any meaningful statistical 
  analysis of the data.  Devon Allen had previously ran the third fastest time in history when 
  he ran 12.84 earlier in 2022 and was considered to be one of the favorites to win.  His 
  disqualification thus prompted questions into whether the timing equipment used at the World
  Track Championships was accurate and whether or not Allen actually reacted as fast as the 
  timing equipment stated he did.

  This paper questions the reaction time disqualification barrier, which is
  currently 0.1 seconds to determine whether this is a reasonable threshold as
  well as look at data that calls into question whether the timing equipment
  used at the 2022 Track and Field World Championships was faulty.
  There is little at this point that could be done to rectify any potential 
  mis-disqualification other than to try to prevent it from happening again.  
  This matter needs to be addressed because Devon Allen's disqulification could 
  be a problem that could continue to re-appear.  If athletes
  are reacting fast enough that the timing system thinks that they are reacting
  illegally fast, then World Athletics needs to address this and possibly change
  the minimum reaction time allowed.

@@

1)
After coming in third in the 110 meter hurdles at the United States Track and 
Field Championships in Eugene Oregon at the end of June, Allen was expected to 
be a contender for the World Championships, which were scheduled for two months 
later, also in Eugene. At the World Championships, Allen made it through the 
heats and semifinals to reach the finals while posting reaction times of 0.123
in the heats and 0.101 in the semifinal.  Allen, who attended the University
of Oregon (Location of US World Championship and the World Championship) was
considered to be a favorite for the 110 meter hurdle after running 12.84 seconds
in early June; only 0.04 seconds away from the world record \citep{Preview}.
Thus his disqualification by reacting in 0.099 seconds instead of 0.01 was met
with shock from both Allen and fans alike.  The crowd erupted with jeers when
they found out that Allen was disqualified, angry at the result.  Allen himself
appeared frustrated and angered with the result. Had he reacted just 0.001 
seconds slower he would have been allowed to compete and been spared waiting
another two years until the next World Track and Field Championship.
%Do I need to cite a video?

2) and 3)
Other scholarly papers have discussed track and field reaction times but not in the
context of whether the minimum reaction time barrier is too low or whether timing
systems were accurate.  Outside the blog posts from LetsRun.com, there have not been
any published papers regarding Devon Allen's disqualification in particular.  This is not
necessarily surprising given that his disqualification occurred only last August.
Letsrun is a website that is part message board and part news.  The message board,
which functions similarly to Reddit is very active during important running
events, which include during the finals of major events such as
the Olympics and the World Track and Field Championships.  However, it is the
news section of the website that generated articles such as "Was Devon Allen
Screwed? There's At Least a 99\% That He Was" and "The Data Keeps Pouring In and
It Continues to Look Bad For World Athletics and Great For Devon Allen".  Robert
Johnson, who created and runs the Letsrun website argued that there was an error
with the timing equipment at the 2022 Track and Field World Championship and
cited graphs, statistics such as median reaction time, and a reference to a study
published in cooperation with IAAF in 2009 questioning the validity of the 0.1 
second reaction barrier.  The paper analyzed sprinters reaction times
and the biological processes that went into reacting to the starting gun and
concluded that athletes can react in as low as 0.08 seconds. It even recommended
to IAAF that the disqualification barrier be decreased to allow sprinters to
react faster, and suggested 0.08 or 0.085 seconds as possible new thresholds.
Additionally, it was recommended to IAAF that high speed cameras be used to
change the reaction time criteria from pushing off on the block to body movement.
This could provide a more accurate representation of reaction time, and thus the
possibility of a false disqualification would be lowered \citep{Komi}.
Despite the study, IAAF has not changed the reaction barrier since it was
implemented at the 1999 World Track and Field Championships.  It is confusing
as to why IAAF would commission a study only to disregard the results and not
make the necessary changes.  It is unknown how many athletes since 2009 could
have potentially benefited from this, but it is probable that there are others
besides Allen who were disqualified.

@@

Another objective of this paper is to compare data for athletes who have competed
at additional competitions to see how their reaction times compare from previous
competitions to this one. First we can look at the reaction times of athletes who competed at the
World Track and Field Championship in both 2019 and 2022 and compare the times,
and secondly we can look at how athletes performed compared to how they did at
their national/qualifiers meet.  We would expect that athletes of this caliber,
who are world-class athletes, to not improve significantly from one year to the
next.  In order to reduce any chances of athletes improving reaction time, we
can also look at data from qualifying meets.  Almost every athlete had to qualify 
by proving they are elite at the national level, usually by coming in at least
fourth. Thus, data was compiled on athletes who competed
at the 2022 World Track and Field Championships on what these athletes also ran 
at their country's national championship. If there was any argument that athletes
have improved their reaction time from 2019 to 2022, it should now be considered
nearly null, as the national competitions took place between June and July 2022 
and the World Championships were in August.  Thus, it should be expected that 
elite athletes such as these will not get reasonably faster in such a short time
span. The equipment used to measure reaction time cannot be guaranteed to be
the same from national to international competition, but the time span between
meets has been reduced considerably.

5)

The contribution of this paper is to question whether or not Seiko's timing data
was accurate and thus if Devon Allen was disqualified more so because of faulty
equipment rather than reacting illegally fast.  Currently the barrier is 0.10 seconds, any time
faster than this would result in a disqualification in the athlete who reacted too
quickly and the race is restarted \citep{Seiko-Timing}.  Since 2011, any athletes who false 
start are not given a second chance, they are removed from competition \citep{False-Start}.  
This can have massive implications as events like the Outdoor World Track and Field Championship 
only occur typically once every two years.  It is an opportunity for athletes to earn endorsement
deals and prove to sponsors such as Nike and Adidas that
they deserve a contract.  Thus the significance of a disqualification, especially in the finals 
of a major event such as the World Track and Field Championships, are very
consequential.

@@

Something that is extremely important to note in the Data is the inconsistency of World Athletics 
in terms of how they penalize false starts.  From 2007-2009, the IAAF (the former name for World 
Athletics, and the governing body for the World Track and Field Champions) instituted a rule change
that allowed one warning false start before a sprinter was disqualified \citep{False-Start}  This 
essentially gave sprinters a second chance and reduced the penalty for a false start.  There were 
18 male and 7 female false starts at the 2007 World Championships, 18 male and 7 female starts at 
the 2009 World Championships. Starting in 2011, the rule was scrapped, and the old policy which 
automatically disqualifies runners who false start was reinstituted.  The effect? At the 2011 World
Championship there were 6 male disqualifications and 4 female disqualifications \citep{False-Start}. 
By returning to the harsher policy and cracking down on false starts, World Athletics had reduced the
number of false starts by two thirds. This was desirable for World Athletics as false starts can 
make already long track meets more tedious, both for athletes and viewers watching the 
television broadcast.  Since 2011, IAAF has instituted the one false start rule
\citep{Pilianidis}.

Data was copied from the World Athletics website and pasted into an Excel
spreadsheet. The data covers the men's 110 meter hurdles and the women's 100
meter hurdles from 1999 to 2022.  The variable of prediction interest is reaction
time, and the predicting variables are year, stage of competition (heats, 
semifinal, final), total time, and gender.  The data can be found in the appendix:.  
Factors such as name, country, and identification number were
not considered important for the purposes of this research besides the times for
Devon Allen and the other United States athletes mentioned above.  The variable "Stage"
refers to whether the observation occurred during the "heats" (Preliminary round) which
is denoted by "H" in the data, "S" for "semifinals" (Of which there are either 2 or 3 
heats each year), and "F" for finals. The variable "Gender" has levels "M" and "F" 
which stand for "male" and "female" respectively.  The variable "Batch" takes a 
numerical value and every heat of every round from 1999 to 2022 was assigned it's
own number.  This was added because of suspected issues that reaction time may be
correlated with other runners reacting quickly (one runner reacting quickly may
result in other runners reacting quickly).  There is no statistical signficance to
how the batches were labeled, they were ordered chronologically by meet but
reverse chronologically by year (Batch 1 is a preliminary Heat of 2022 and Batch
111 corresponds to the 1999 finals). The data that looked at in the paper
is best visualized using a box plot. Figure~\ref{fig:DataBoxPlot}.  Included are
box plots of the reaction times of the men's semi finals and finals pooled
together and just the finals.

\begin{figure}[h]
  \centering 
  \includegraphics{DataBoxPlot}
  \label{fig:DataBoxPlot}
\end{figure}

@@

\subsection{GLMM}
A gamma mixed linear effects model was chosen as the best model to model the data.
Many models were tried and compared to determine what the strongest model.  For
a fixed data set such as the men's pooled data including 2022, the models with
the venue effect included performed better than the models without the venue
effect for each set of comparable models; as indicated by the lower AIC and
higher log likelihood.  Thus, the venue effect is significant and should be
included in the model.  It can also be established that a gamma distribution
better models the data than linear data from comparing the linear model with
venue effect and them gamma model with a venue effect.  The final decision that
was made in model selection was to include link = log in the model.  Despite
%Does this part need to be changed?
doing very little to alter the AIC or log likelihood, the log scale was included
in order to try and better deal with high outliers. %Include Residuals plot?

All Men's semi finals and finals data
\begin{center}
  \begin{tabular}{|c  c  c  c  c  c  c |}
   \hline 
   Model & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
   \hline\hline
   linear one & -1667 & 835.5 & 0.1529 & 0.0014 & N/A & N/A \\
   \hline
   linear year & -1713 & 859.3 & 0.1541 & 0.0040 & 1.719e-04 & 5.926e-04 \\ 
   \hline
   gamma one & -1756 & 880.2 & 6.554 & 0.0598 & N/A & N/A \\
   \hline
   gamma year & -1848 & 926.9 & 6.553 & 0.2269 & 0.1305 & 0.0234 \\
   \hline
   gamma one log & -1756 & 880.2 & -1.880 & 0.0091 & N/A & N/A \\
   \hline
   gamma year log & -1848 & 927.0 & -1.869 & 0.0348 & 0.0031 & 0.0235 \\ [0.5ex]
   \hline
  \end{tabular}
  \end{center}

@@

\subsection{Rank-based comparison for Clustered}
This section was motivated from a qualitative analysis that started by looking
at how United States athletes reacted at the 2022 United States Track and Field
Championships. From June 23-26 2022, USA held its Track
and Field Championship to decide who to send to the World Championships at Hayward 
Field, the exact same venue as used for the World Track and Field Championship a month 
later.  Thus, a baseline is able to be established for the four USA athletes who 
competed in at least one round of both events: Trey Cunningham, Daniel Roberts, 
Grant Holloway, and Devon Allen. Every athlete reacted faster in all of the World 
Track and Field Championship races compared to the USA Track and Field 
Championship races. Devon Allen raced three times at the USA Track and
Field Championship and three times at the World Track and Field  Championship. 
His reaction times were significantly faster in the World Track and
Field Championship compared to the USA Track and Field Championship as were his
teamates.  This comparison suggests that the timing system used between the two
meets may have been meaningfully different to the point where Devon Allen was
disqualified more because of faulty equipment rather than because he reacted
too quickly.  This table below highlights the difference in reaction times for
Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. "USA" denotes
the USA Track and Field Championships, "World" denotes the World Track and Field
Championships, "N/A" was used for any athlete that did not compete and thus
did not register a reaction time.  All numbers listed in the table are the reaction
time for the respective athletes in seconds.

\begin{center}
  \begin{tabular}{||c | c c c | c c c||} 
   \hline
   Athlete & USA H & USA S & USA F & World H & World S & World F \\ [0.5ex] 
   \hline\hline
   Devon Allen & 0.201 & 0.153 & 0.160 & 0.123 & 0.101 & 0.099 \\ 
   \hline
   Trey Cunningham & 0.186 & 0.185 & 0.182 & 0.115 & 0.120 & 0.109 \\
   \hline
   Grant Holloway & 0.192 & 0.190 & N/A & 0.147 & 0.128 & 0.124 \\
   \hline
   Daniel Roberts & 0.181 & 0.200 & 0.183 & 0.179 & N/A & N/A \\ [0.5ex]
   \hline
  \end{tabular}
  \end{center}

@@

%Citations needed
Using the "clusWilcox.test" function, a grouped and clustered Wilcoxon rank sum
test was performed wtih the cluster set to be the Athlete and group set to Year.
The method was set to be "ds" which is used for the Data and Satten method, which
is preferable when the cluster sizes are informative \citep{Clusrank}.  We do not care that some
athletes made it further in either competition than another or that it should
not matter how elite an athlete is at the 110 meter hurdles once we get to the
national international competition level.  Cluster sizes vary from 2 observations
to 6.

@@

To try and approximate the distribution of the reaction times, R code was written
to try and estimate the probabilities of observing a reaction time as extreme as
0.1 seconds. At a size of n=12,000,000, the calculated probability of a reaction
time being below 0.1 seconds was as follows for the various different data sets:

@@

>>>>

%Abstract should be 200 words
%Devon Allen was disqualified in 2022... discussions about dq, talk methods and 
%models, venue effect, and conclusion
\begin{abstract}
  At the 2022 World Track Championships in Eugene Oregon, Devon Allen was disqualified in the
  Men's 110 meter hurdle final after registering a reaction time of 0.099 seconds, 0.001 
  seconds faster than what is allowed.  Following the games, bloggers on the running website 
  LetsRun scrutinized the results and performed non-statistical analysis of the data.
  They found that the reaction time data from the 2022 World Track Championships seemed to be faster 
  compared to the other datasets they looked at, but did not perform any meaningful statistical 
  analysis of the data.  Devon Allen had previously run the third-fastest time in history when 
  he ran 12.84 earlier in 2022 and was considered to be one of the favorites to win.  His 
  disqualification thus prompted questions into whether the timing equipment used at the World
  Track Championships was accurate and whether Allen actually reacted as fast as the 
  timing equipment stated he did. This paper questions the reaction time disqualification barrier, which is
  currently 0.1 seconds to determine whether this is a reasonable threshold as
  well as look at data that calls into question whether the timing equipment
  used at the 2022 Track and Field World Championships was faulty.
  There is little at this point that could be done to rectify any potential 
  mis-disqualification other than to try to prevent it from happening again.  
  This matter needs to be addressed because Devon Allen's disqualification could 
  be a problem that could continue to re-appear.  If athletes
  are reacting fast enough that the timing system thinks that they are reacting
  illegally fast, then World Athletics needs to address this and possibly change
  the minimum reaction time allowed.

@@

1)
After coming in third in the 110 meter hurdles at the United States Track and 
Field Championships in Eugene Oregon at the end of June, Allen was expected to 
be a contender for the World Championships, which were scheduled for two months 
later, also in Eugene. At the World Championships, Allen made it through the 
heats and semifinals to reach the finals while posting reaction times of 0.123
in the heats and 0.101 in the semifinal.  Allen, who attended the University
of Oregon (Location of US World Championship and the World Championship) was
considered to be a favorite for the 110 meter hurdle after running 12.84 seconds
in early June; only 0.04 seconds away from the world record \citep{Preview}.
Thus his disqualification by reacting in 0.099 seconds instead of 0.01 was met
with shock from both Allen and fans alike.  The crowd erupted with jeers when
they found out that Allen was disqualified, angry at the result.  Allen himself
appeared frustrated and angered with the result. Had he reacted just 0.001 
seconds slower he would have been allowed to compete and been spared waiting
another two years until the next World Track and Field Championship.
%Can talk about youtube video and put the link in text here

2) and 3)
Other scholarly papers have discussed track and field reaction times but not in the
context of whether the minimum reaction time barrier is too low or whether timing
systems were accurate.  Outside the blog posts from LetsRun.com, there have not been
any published papers regarding Devon Allen's disqualification in particular.  This is not
necessarily surprising given that his disqualification occurred only last August.
Letsrun is a website that is part message board and part news.  The message board,
which functions similarly to Reddit is very active during important running
events, which include during the finals of major events such as
the Olympics and the World Track and Field Championships.  However, it is the
news section of the website that generated articles such as "Was Devon Allen
Screwed? There's At Least a 99\% That He Was" and "The Data Keeps Pouring In and
It Continues to Look Bad For World Athletics and Great For Devon Allen".  Robert
Johnson, who created and runs the Letsrun website argued that there was an error
with the timing equipment at the 2022 Track and Field World Championship and
cited graphs, statistics such as median reaction time, and a reference to a study
published in cooperation with IAAF in 2009 questioning the validity of the 0.1 
second reaction barrier.  The paper analyzed sprinters reaction times
and the biological processes that went into reacting to the starting gun and
concluded that athletes can react in as low as 0.08 seconds. It even recommended
to IAAF that the disqualification barrier be decreased to allow sprinters to
react faster, and suggested 0.08 or 0.085 seconds as possible new thresholds.
Additionally, it was recommended to IAAF that high speed cameras be used to
change the reaction time criteria from pushing off on the block to body movement.
This could provide a more accurate representation of reaction time, and thus the
possibility of a false disqualification would be lowered \citep{Komi}.
Despite the study, IAAF has not changed the reaction barrier since it was
implemented at the 1999 World Track and Field Championships.  It is confusing
as to why IAAF would commission a study only to disregard the results and not
make the necessary changes.  It is unknown how many athletes since 2009 could
have potentially benefited from this, but it is probable that there are others
besides Allen who were disqualified.
%Reaction time barrier has been examined before: cite reviews
%Then go into specifics and talk about people who wanted to change barrier
%General reviews are important
%People have studied the biological processes, cite a few references
%How barrier is picked by regulator
%Read abstract and into and summarize into about one sentence what each paper contributes
%1 and 2 can be merged but 3 need to stand alone

@@

Another objective of this paper is to compare data for athletes who have competed
at additional competitions to see how their reaction times compare from previous
competitions to this one. First we can look at the reaction times of athletes who competed at the
World Track and Field Championship in both 2019 and 2022 and compare the times,
and secondly we can look at how athletes performed compared to how they did at
their national/qualifiers meet.  We would expect that athletes of this caliber,
who are world-class athletes, to not improve significantly from one year to the
next.  In order to reduce any chances of athletes improving reaction time, we
can also look at data from qualifying meets.  Almost every athlete had to qualify 
by proving they are elite at the national level, usually by coming in at least
fourth. Thus, data was compiled on athletes who competed
at the 2022 World Track and Field Championships on what these athletes also ran 
at their country's national championship. If there was any argument that athletes
have improved their reaction time from 2019 to 2022, it should now be considered
nearly null, as the national competitions took place between June and July 2022 
and the World Championships were in August.  Thus, it should be expected that 
elite athletes such as these will not get reasonably faster in such a short time
span. The equipment used to measure reaction time cannot be guaranteed to be
the same from national to international competition, but the time span between
meets has been reduced considerably.
%Simplify contribution to one paragraph: collect data and formal statistical
%analysis such as model used
%Move a lot of this to methods section
5)
%Follow Anthony's work, explicit references as to what each section does

%Seiko section should go in the Results section.  From results it seems reasonable
%question the data and talk about Seiko
The contribution of this paper is to question whether or not Seiko's timing data
was accurate and thus if Devon Allen was disqualified more so because of faulty
equipment rather than reacting illegally fast.  Currently the barrier is 0.10 seconds, any time
faster than this would result in a disqualification in the athlete who reacted too
quickly and the race is restarted \citep{Seiko-Timing}.  Since 2011, any athletes who false 
start are not given a second chance, they are removed from competition \citep{False-Start}.  
This can have massive implications as events like the Outdoor World Track and Field Championship 
only occur typically once every two years.  It is an opportunity for athletes to earn endorsement
deals and prove to sponsors such as Nike and Adidas that
they deserve a contract.  Thus the significance of a disqualification, especially in the finals 
of a major event such as the World Track and Field Championships, are very
consequential.

@@

  %yij denotes ith years and j is observation in year
  %yij will have clustered i and we use random effect to account for this
  %When model is presented talk about how I modeled yij
  %Zij where player is i and cluster, group is year and j
  %Test hypothesis that reaction times are the same vs that they are not
Something that is extremely important to note in the Data is the inconsistency of World Athletics 
in terms of how they penalize false starts.  From 2007-2009, the IAAF (the former name for World 
Athletics, and the governing body for the World Track and Field Champions) instituted a rule change
that allowed one warning false start before a sprinter was disqualified \citep{False-Start}  This 
essentially gave sprinters a second chance and reduced the penalty for a false start.  There were 
18 male and 7 female false starts at the 2007 World Championships, 18 male and 7 female starts at 
the 2009 World Championships. Starting in 2011, the rule was scrapped, and the old policy which 
automatically disqualifies runners who false start was reinstituted.  The effect? At the 2011 World
Championship there were 6 male disqualifications and 4 female disqualifications \citep{False-Start}. 
By returning to the harsher policy and cracking down on false starts, World Athletics had reduced the
number of false starts by two thirds. This was desirable for World Athletics as false starts can 
make already long track meets more tedious, both for athletes and viewers watching the 
television broadcast.  Since 2011, IAAF has instituted the one false start rule
\citep{Pilianidis}.

Data was copied from the World Athletics website and pasted into an Excel
spreadsheet. The data covers the men's 110 meter hurdles and the women's 100
meter hurdles from 1999 to 2022.  The variable of prediction interest is reaction
time, and the predicting variables are year, stage of competition (heats, 
semifinal, final), total time, and gender.  The data can be found in the appendix:.  
Factors such as name, country, and identification number were
not considered important for the purposes of this research besides the times for
Devon Allen and the other United States athletes mentioned above.  The variable "Stage"
refers to whether the observation occurred during the "heats" (Preliminary round) which
is denoted by "H" in the data, "S" for "semifinals" (Of which there are either 2 or 3 
heats each year), and "F" for finals. The variable "Gender" has levels "M" and "F" 
which stand for "male" and "female" respectively.  The variable "Batch" takes a 
numerical value and every heat of every round from 1999 to 2022 was assigned it's
own number.  This was added because of suspected issues that reaction time may be
correlated with other runners reacting quickly (one runner reacting quickly may
result in other runners reacting quickly).  There is no statistical signficance to
how the batches were labeled, they were ordered chronologically by meet but
reverse chronologically by year (Batch 1 is a preliminary Heat of 2022 and Batch
111 corresponds to the 1999 finals). The data that looked at in the paper
is best visualized using a box plot. Figure~\ref{fig:DataBoxPlot}.  Included are
box plots of the reaction times of the men's semi finals and finals pooled
together and just the finals.
%Fix quotations
%Dont need the titles will add captions, use width =5, height = 3 in aspect ratio
\begin{figure}[h]
  \centering 
  \includegraphics{DataBoxPlot}
  \label{fig:DataBoxPlot}
\end{figure}

@@

\subsection{GLMM}
%Include latex math symbols for the equations
A gamma mixed linear effects model was chosen as the best model to model the data.
Many models were tried and compared to determine what the strongest model.  For
a fixed data set such as the men's pooled data including 2022, the models with
the venue effect included performed better than the models without the venue
effect for each set of comparable models; as indicated by the lower AIC and
higher log likelihood.  Thus, the venue effect is significant and should be
included in the model.  It can also be established that a gamma distribution
better models the data than linear data from comparing the linear model with
venue effect and them gamma model with a venue effect.  The final decision that
was made in model selection was to include link = log in the model.  Despite
%Does this part need to be changed?
doing very little to alter the AIC or log likelihood, the log scale was included
in order to try and better deal with high outliers. %Include Residuals plot?

%Throughout exploratory analysis, gamma provides better AIC
%Spend paragraph on after fitting the model how to figure out 0.1 p value

%Table should go to the results section
All Men's semi finals and finals data
\begin{center}
  \begin{tabular}{|c  c  c  c  c  c  c |}
   \hline 
   Model & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
   \hline\hline
   linear one & -1667 & 835.5 & 0.1529 & 0.0014 & N/A & N/A \\
   \hline
   linear year & -1713 & 859.3 & 0.1541 & 0.0040 & 1.719e-04 & 5.926e-04 \\ 
   \hline
   gamma one & -1756 & 880.2 & 6.554 & 0.0598 & N/A & N/A \\
   \hline
   gamma year & -1848 & 926.9 & 6.553 & 0.2269 & 0.1305 & 0.0234 \\
   \hline
   gamma one log & -1756 & 880.2 & -1.880 & 0.0091 & N/A & N/A \\
   \hline
   gamma year log & -1848 & 927.0 & -1.869 & 0.0348 & 0.0031 & 0.0235 \\ [0.5ex]
   \hline
  \end{tabular}
  \end{center}

@@

\subsection{Rank-based comparison for Clustered}
%This should go to data section
This section was motivated from a qualitative analysis that started by looking
at how United States athletes reacted at the 2022 United States Track and Field
Championships. From June 23-26 2022, USA held its Track
and Field Championship to decide who to send to the World Championships at Hayward 
Field, the exact same venue as used for the World Track and Field Championship a month 
later.  Thus, a baseline is able to be established for the four USA athletes who 
competed in at least one round of both events: Trey Cunningham, Daniel Roberts, 
Grant Holloway, and Devon Allen. Every athlete reacted faster in all of the World 
Track and Field Championship races compared to the USA Track and Field 
Championship races. Devon Allen raced three times at the USA Track and
Field Championship and three times at the World Track and Field  Championship. 
His reaction times were significantly faster in the World Track and
Field Championship compared to the USA Track and Field Championship as were his
teamates.  This comparison suggests that the timing system used between the two
meets may have been meaningfully different to the point where Devon Allen was
disqualified more because of faulty equipment rather than because he reacted
too quickly.  This table below highlights the difference in reaction times for
Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. "USA" denotes
the USA Track and Field Championships, "World" denotes the World Track and Field
Championships, "N/A" was used for any athlete that did not compete and thus
did not register a reaction time.  All numbers listed in the table are the reaction
time for the respective athletes in seconds.

\begin{center}
  \begin{tabular}{|c c c c c c c|} 
   \hline
   Athlete & USA H & USA S & USA F & World H & World S & World F \\ [0.5ex] 
   \hline\hline
   Devon Allen & 0.201 & 0.153 & 0.160 & 0.123 & 0.101 & 0.099 \\ 
   \hline
   Trey Cunningham & 0.186 & 0.185 & 0.182 & 0.115 & 0.120 & 0.109 \\
   \hline
   Grant Holloway & 0.192 & 0.190 & N/A & 0.147 & 0.128 & 0.124 \\
   \hline
   Daniel Roberts & 0.181 & 0.200 & 0.183 & 0.179 & N/A & N/A \\ [0.5ex]
   \hline
  \end{tabular}
  \end{center}

@@

%Citations needed
%Include math formulas and the test statistic
%Equation of test statistic is from papers cited by clusrank. Asymptootic dist
%Was found by __ referecne and the implementation is from the clusrank R package
Using the "clusWilcox.test" function, a grouped and clustered Wilcoxon rank sum
test was performed with the cluster set to be the Athlete and group set to Year.
The method was set to be "ds" which is used for the Data and Satten method, which
is preferable when the cluster sizes are informative \citep{Clusrank}.  We do not care that some
athletes made it further in either competition than another or that it should
not matter how elite an athlete is at the 110 meter hurdles once we get to the
national international competition level.  Cluster sizes vary from 2 observations
to 6.

@@

%This should go in the method section, generate from fitted model explain what
%The R code does.  Simulate from fitted model. Use distribution of model to 
%Figure out the p value of reaction time less than 0.1
To try and approximate the distribution of the reaction times, R code was written
to try and estimate the probabilities of observing a reaction time as extreme as
0.1 seconds. At a size of n=12,000,000, the calculated probability of a reaction
time being below 0.1 seconds was as follows for the various different data sets:

@@

%Section 4.1: Gamma model results show the fitting with and without 2022,
%talk about p values and thresholds from the simulations (do not include the
%results in the methods section).  Discussion should show that random effect
%and standard error increase and that 2022 is special.  Now talk about Seiko
%timing equipment

%Section 4.2 is motivated by seiko timing and look at same athletes across
%multiple types of competition: 2019 and national

%Discussion/Conclusion is beyond results.  Limitations, what type of data we
%hope to have.
